---
title: "Grade 5 Reading"
author: "Logan Brady"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Importing library packages
```{r}
library(arsenal)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(tidyverse)
library(data.table)
library(readxl)
library(readr)
```

#Importing th Grade Reading Data files
```{r}

Reading_Fall5 <- read_excel("C:/Users/Logan/OneDrive/Desktop/VGA Analysis/Raw VGA Scores/5 VGA analysis 22-23.xlsx", sheet = 1)
Reading_Winter5 <- read_excel("C:/Users/Logan/OneDrive/Desktop/VGA Analysis/Raw VGA Scores/5 VGA analysis 22-23.xlsx", sheet = 2)

View(Reading_Fall5)
View(Reading_Winter5)
```

#Cleaning the datasets.

#Comparing two datasets using comparedf()
```{r}
summary(comparedf(Reading_Fall5, Reading_Winter5))
```

#How to return column titles
```{r}
head(x = Reading_Fall5)
colnames(x = Reading_Fall5)

head(x = Reading_Winter5)
colnames(x = Reading_Winter5)
```

#Rename data columns
```{r}
colnames(Reading_Fall5)[1] = "Last Name"
colnames(Reading_Fall5)[2] = "First Name"
colnames(Reading_Fall5)[3] = "Teacher"
colnames(Reading_Fall5)[4] = "Vertical Scaled Score"

colnames(Reading_Winter5)[1] = "Last Name"
colnames(Reading_Winter5)[2] = "First Name"
colnames(Reading_Winter5)[3] = "Teacher"
colnames(Reading_Winter5)[4] = "Vertical Scaled Score"

colnames(x = Reading_Fall5)
colnames(x = Reading_Winter5)
```

#Manipulating record numbers 61, 89, 90, 97, 112  in Reading_Fall5 to make them unique indentifiers. Changing "LONG" to "LONG S.", "RODGERS" to "RODGERS W.", "RODRIGUEZ" to "RODRIGUEZ J.", "SAYLOR" to "SAYLOR A.", and "VARGAS" to "VARGAS N.".
```{r}
Reading_Fall5[61, 1] <- "LONG S."
Reading_Fall5[89, 1] <- "RODGERS W."
Reading_Fall5[90, 1] <- "RODRIGUEZ J."
Reading_Fall5[97, 1] <- "SAYLOR A."
Reading_Fall5[112, 1] <- "VARGAS N."

view(Reading_Fall5)
```

