---
title: "VGA Data Cleaning"
author: "Logan Brady"
date: "2023-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#DATA CLEANING

#Import library packages
```{r}
library(readr)
library(data.table)
library(dplyr)
library(ggpubr)
```

#Import data file
```{r}
VGAdata <- read_csv("C:/Users/Logan/OneDrive/Desktop/VGA Analysis/Raw VGA Scores/Winter VGA Scores.csv")
View(VGAdata)
```

#Return top portion of the dataset.
```{r}
head(x = VGAdata)
```

#Return column titles
```{r}
colnames(x = VGAdata)
```


#Remove ALL FILLER COLUMNS.

#How to remove any "filler" columns from the dataset.
```{r}
VGAdata[ , c("Filler...7", "Filler...11", "Filler...22", "Filler...25", "Filler...34", "Filler...35", "Filler...37", "Filler...40", "Filler...41", "Filler...42", "Filler...47", "Filler...48", "Filler...49", "Filler...50", "Filler...51", "Filler...52", "Filler...53", "Filler...54", "Filler...55", "Filler...57", "Filler...59", "Filler...62", "Filler...63", "Filler...64", "Filler...65", "Filler...66", "Filler...67", "Filler...71", "Filler...73", "Filler...74", "Filler...76", "Filler...86", "Filler...91", "Filler...93", "Filler...95", "Filler...96", "Filler...97", "Filler...98", "Filler...99", "Filler...100", "Filler...101", "Filler...111", "Filler...113", "Filler...115", "Filler...116", "Filler...117", "Filler...126", "Filler...127", "Filler...128", "Filler...152", "Filler...153", "Filler...154", "Filler...159", "Filler...162")] <- list(NULL)

colnames(x = VGAdata)
```


#Remove variables that do not have any useful data.

#How to remove columns "STA-B", "Code VTLN", "Code TLN", "Code TFN", "Performance Level", "Reporting Cat Scaled Score 1", "Reporting Cat Scaled Score 2", "Reporting Cat Scaled Score 3", "Reporting Cat Scaled Score 4", "Reporting Cat Scaled Score 5", "Reporting Cat Scaled Score 6", "Reporting Cat Scaled Score 7", "Testing Status", "Z-Code-G", "Z-Code-F", "Z-Code-E", "SLIFE", "Recovery", "Retest", "Recently Arrived EL", "X-Code-D", "X-Code-c", "X-Code-B", "Foster Care", "Local Use Data - Student", "Virtual Virginia", "Temp Condition", "Resident Division", "Transfer", "SOA Adjustment - EL", "Middle Initial", and "Group Code" from the dataset.
```{r}
VGAdata[ , c("Z indicates End of Record", "STA-B", "Code VTLN", "Code TLN", "Code TFN", "Performance Level", "Reporting Cat Scaled Score 1", 
             "Reporting Cat Scaled Score 2", "Reporting Cat Scaled Score 3", "Reporting Cat Scaled Score 4", "Reporting Cat Scaled Score 5", 
             "Reporting Cat Scaled Score 6", "Reporting Cat Scaled Score 7", "Test Scaled Score", "Testing Status", "Z-Code-G", "Z-Code-F", "Z-Code-E", "SLIFE", 
             "Recovery", "Retest", "Recently Arrived EL", "X-Code-D", "X-Code-C", "X-Code-B", "Foster Care", "Local Use Data - Student", "Virtual Virginia",
             "Temp Condition", "Resident Division", "Transfer", "SOA Adjustment - EL", "Middle Initial", "Group Code", "Student Test Day", "Form", "Level", "Core/Admin", "Year of Standard",              "Test Administration-Time of Year", "Test Administration-Year", "Sequence Number", "Document ID", "Batch Number")] <- list(NULL)

colnames(x = VGAdata)
```

#Manipulate the values in Disability Status from NA to 0 and any value greater than 0 to 1. The type of disability isn't necessary for this form of analysis.
```{r}
VGAdata["Disability Status"][is.na(VGAdata["Disability Status"])] <- 0

View(VGAdata)
```

#Manipulate the values in  from NA to 0.
```{r}
VGAdata["Primary Nighttime Residence Code"][is.na(VGAdata["Primary Nighttime Residence Code"])] <- 0
View(VGAdata)
```

#Manipulate the values in Formerly EL from NA to 0 and any value greater than 0 to 1. Just need to verify who is EL and who's not.
```{r}
VGAdata["Formerly EL"][is.na(VGAdata["Formerly EL"])] <- 0
VGAdata$"Formerly EL"[VGAdata$"Formerly EL" > 0] <- 1

View(VGAdata)
```

#Remove EL Overall Proficiency Level
```{r}
VGAdata[ , c("EL Overall Proficiency Level")] <- list(NULL)
View(VGAdata)
```

#Manipulate the values in Stack Number, all NA to 0.
```{r}
VGAdata["Stack Number"][is.na(VGAdata["Stack Number"])] <- 0
View(VGAdata)
```

#Manipulate the values in Revision, all 0 to 1 and NA to 0.
```{r}
VGAdata["Revision"][VGAdata["Revision"] == 0] <- 1
VGAdata["Revision"][is.na(VGAdata["Revision"])] <- 0

View(VGAdata)
```

#Manipulate the values in , all 0 to 1 and NA to 0.
```{r}
VGAdata["Items Blank &/or Mult Marked"][VGAdata["Items Blank &/or Mult Marked"] == 0] <- 1
VGAdata["Items Blank &/or Mult Marked"][is.na(VGAdata["Items Blank &/or Mult Marked"])] <- 0

View(VGAdata)
```

#Manipulate the test name for the student with test revisions.
```{r}
VGAdata["Test Name"][VGAdata["Test Name"] == "4 Math Growth"] <- "4 Math Growth CAT"
VGAdata["Test Name"][VGAdata["Test Name"] == "4 Read Growth"] <- "4 Read Growth CAT"

View(VGAdata)
```

#Manipulate the test name to just Read/Math Growth CAT
```{r}
VGAdata["Test Name"][VGAdata["Test Name"] == "3 Math Growth CAT"] <- "Math Growth CAT"
VGAdata["Test Name"][VGAdata["Test Name"] == "3 Read Growth CAT"] <- "Read Growth CAT"

VGAdata["Test Name"][VGAdata["Test Name"] == "4 Math Growth CAT"] <- "Math Growth CAT"
VGAdata["Test Name"][VGAdata["Test Name"] == "4 Read Growth CAT"] <- "Read Growth CAT"

VGAdata["Test Name"][VGAdata["Test Name"] == "5 Math Growth CAT"] <- "Math Growth CAT"
VGAdata["Test Name"][VGAdata["Test Name"] == "5 Read Growth CAT"] <- "Read Growth CAT"

View(VGAdata)
```

#Find remaining NA Values
```{r}
which(colSums(is.na(VGAdata)) > 0)
```


#Manipulate Arya Williams data to update student number to 697550 from NA. As well as her "Local Use-Test" from NA to SRI.
```{r}
VGAdata["Student Number"][is.na(VGAdata["Student Number"])] <- 697550
VGAdata["Local Use Data -Test"][is.na(VGAdata["Local Use Data -Test"])] <- "SRI"
```

#Manipulate the NA values in Online Test to N to show that it was a revised paper test.
```{r}
VGAdata["Online Test"][is.na(VGAdata["Online Test"])] <- "N"
```

#Manipulate the STA-Codes in the datasheet. Where it says NA change it to N.
```{r}
VGAdata[is.na(VGAdata)] <- "N"
```

#Write VGAdata as CSV file for PowerBI.
```{r}
write.csv(VGAdata, "NewestVGAdata.csv")
```



#FORMERLY_EL, importing file named "FormerlyEL" from folder named "Useable Files" which were created in the "Initial Raw VGA Data Cleaning".
```{r}
library(readr)
FormerlyEL <- read_csv("Useable Files/FormerlyEL.csv")
View(FormerlyEL)
```
#Return column names and those with NA values.
```{r}
colnames(x = FormerlyEL)
which(colSums(is.na(FormerlyEL)) > 0)
```

#Delete columns with NA Values.
```{r}
FormerlyEL[ , c("Primary Nighttime Residence Code", "EL Overall Proficiency Level", "Stack Number", "Batch Number", 
                "Sequence Number", "Document ID", "Revision", "Items Blank &/or Mult Marked")] <- list(NULL)
View(FormerlyEL)
```

#Manipulate the test name to just Read/Math Growth CAT
```{r}
FormerlyEL["Test Name"][FormerlyEL["Test Name"] == "3 Math Growth CAT"] <- "Math Growth CAT"
FormerlyEL["Test Name"][FormerlyEL["Test Name"] == "3 Read Growth CAT"] <- "Read Growth CAT"

FormerlyEL["Test Name"][FormerlyEL["Test Name"] == "4 Math Growth CAT"] <- "Math Growth CAT"
FormerlyEL["Test Name"][FormerlyEL["Test Name"] == "4 Read Growth CAT"] <- "Read Growth CAT"

FormerlyEL["Test Name"][FormerlyEL["Test Name"] == "5 Math Growth CAT"] <- "Math Growth CAT"
FormerlyEL["Test Name"][FormerlyEL["Test Name"] == "5 Read Growth CAT"] <- "Read Growth CAT"

View(FormerlyEL)
```

#Removing Columns that aren't needed.
```{r}
FormerlyEL[ , c("Local Use Data -Test", "Test Administration-Time of Year", "Test Administration-Year", "Stack Number", "Batch Number", "Sequence Number", "Document ID", "Level", "Form", "Subject Code", "Core/Admin", "Year of Standard", "UIN", "UUID", "Z indicates End of Record")] <- list(NULL)
View(FormerlyEL)

```

#Save as CSV file
```{r}
write.csv(FormerlyEL, "NewestFormerlyEL.csv")
```





#EL_LEVEL DATA CLEANING


```{r}
EL_Level <- read_csv("Useable FIles/EL_Level.csv")
View(EL_Level)
```

#Return column names and those with NA values.
```{r}
colnames(x = EL_Level)
which(colSums(is.na(EL_Level)) > 0)
```

#Delete columns with NA Values.
```{r}
EL_Level[ , c("Primary Nighttime Residence Code", "Stack Number", "Batch Number", "Sequence Number", 
                "Document ID", "Revision", "Formerly EL", "Items Blank &/or Mult Marked")] <- list(NULL)
View(EL_Level)
```

#Manipulate the test name to just Read/Math Growth CAT
```{r}
EL_Level["Test Name"][EL_Level["Test Name"] == "3 Math Growth CAT"] <- "Math Growth CAT"
EL_Level["Test Name"][EL_Level["Test Name"] == "3 Read Growth CAT"] <- "Read Growth CAT"

EL_Level["Test Name"][EL_Level["Test Name"] == "4 Math Growth CAT"] <- "Math Growth CAT"
EL_Level["Test Name"][EL_Level["Test Name"] == "4 Read Growth CAT"] <- "Read Growth CAT"

EL_Level["Test Name"][EL_Level["Test Name"] == "5 Math Growth CAT"] <- "Math Growth CAT"
EL_Level["Test Name"][EL_Level["Test Name"] == "5 Read Growth CAT"] <- "Read Growth CAT"

View(EL_Level)
```

#Removing Columns that aren't needed.
```{r}
EL_Level[ , c("Division Name", "Division Code", "School Name", "School Code", "Local Use Data -Test", "Test Administration-Time of Year", 
              "Test Administration-Year", "Level", "Form", "Subject Code", "Core/Admin", "Year of Standard", "UIN", "Online Test", "UUID", 
              "Z indicates End of Record")] <- list(NULL)
View(EL_Level)

```

#Find remaining NA Values
```{r}
which(colSums(is.na(EL_Level)) > 0)
```
#Converting Group Name Values to represent teacher last names.
```{r}
unique(EL_Level$"Group Name")
```

#Writing the code to convert the Group Name values.
```{r}
EL_Level["Group Name"][EL_Level["Group Name"] == "FA PerryS"] <- "Perry"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI PerryS"] <- "Perry"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA JardelR"] <- "Jardel"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI JardelR"] <- "Jardel"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA FlisJ"] <- "Flis"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA BojarczukC"] <- "Bojarczuk"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI BojarczukC"] <- "Bojarczuk"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI SteenK"] <- "Steen"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI JohnsonW"] <- "Johnson"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA ThriftM"] <- "Thrift"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI ThriftM"] <- "Thrift"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA ConlanS"] <- "Conlan"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI ConlanS"] <- "Conlan"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA GrantB"] <- "Grant"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI GrantB"] <- "Grant"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA BreuerE"] <- "Breuer"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI BreuerE"] <- "Breuer"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA CampbellS"] <- "Campbell"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI CampbellS"] <- "Campbell"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA PeoplesN"] <- "Peoples"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI PeoplesN"] <- "Peoples"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA WittM"] <- "Witt"
EL_Level["Group Name"][EL_Level["Group Name"] == "FA FickT"] <- "Fick"
EL_Level["Group Name"][EL_Level["Group Name"] == "WI FickT"] <- "Fick"

View(EL_Level)
```


#Save as CSV file
```{r}
write.csv(EL_Level, "NewestEL_Level.csv")
```




#Disability Status

#Return records of students who have a recorded disability.
```{r}
Disability_Status <- VGAdata[VGAdata$"Disability Status" > 0, ]
View(Disability_Status)
write.csv(Disability_Status, "Disability_Status.csv")
```

#Converting Group Name Values to represent teacher last names.
```{r}
unique(Disability_Status$"Group Name")
```

#Writing the code to convert the Group Name values.
```{r}
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA SteenK"] <- "Steen"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI SteenK"] <- "Steen"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA JohnsonW"] <- "Johnson"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI JohnsonW"] <- "Johnson"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA BojarczukC"] <- "Bojarczuk"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI BojarczukC"] <- "Bojarczuk"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA JardelR"] <- "Jardel"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI JardelR"] <- "Jardel"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA PerryS"] <- "Perry"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI PerryS"] <- "Perry"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA FlisJ"] <- "Flis"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA ThorpK"] <- "Thorp"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI ThorpK"] <- "Thorp"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA GrantB"] <- "Grant"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI GrantB"] <- "Grant"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA ConlanS"] <- "Conlan"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI ConlanS"] <- "Conlan"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA PeoplesN"] <- "Peoples"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI PeoplesN"] <- "Peoples"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA ThriftM"] <- "Thrift"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI ThriftM"] <- "Thrift"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA FickT"] <- "Fick"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI FickT"] <- "Fick"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA MozingoS"] <- "Mozingo"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI MozingoS"] <- "Mozingo"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA RimkusJ"] <- "Rimkus"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI RimkusJ"] <- "Rimkus"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA WittM"] <- "Witt"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "FA CampbellS"] <- "Campbell"
Disability_Status["Group Name"][Disability_Status["Group Name"] == "WI CampbellS"] <- "Campbell"


View(Disability_Status)
```

#Return column names.
```{r}
colnames(x = Disability_Status)
```

#Find remaining NA Values
```{r}
which(colSums(is.na(Disability_Status)) > 0)
```

#Removing Columns that aren't needed.
```{r}
Disability_Status[ , c("Division Name", "Division Code", "School Name", "School Code", "Local Use Data -Test", "Test Administration-Time of Year", 
              "Test Administration-Year", "Level", "Form", "Subject Code", "Core/Admin", "Year of Standard", "UIN", "Online Test", "UUID", 
              "Z indicates End of Record")] <- list(NULL)
View(Disability_Status)

```

#Save as CSV file
```{r}
write.csv(Disability_Status, "Disability_Status.csv")
```


#Correlation coefficient analysis for EL_Level. THIS IS A GOOD MODEL!!!
```{r}
ggscatter(EL_Level,
          x = "EL Overall Proficiency Level",
          y = "Vertical Scaled Score",
          add = "reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          cor.method = "pearson",
          xlab = "ESL Proficiency Level",
          ylab = "Vertical Scaled Score")
```

#Correlation coefficient analysis for VGAdata. Checking for possible racial underrepresentation! NOT A GOOD MODEL!!!
```{r}
ggscatter(VGAdata,
          x = "Race",
          y = "Vertical Scaled Score",
          add = "reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          cor.method = "pearson",
          xlab = "Race",
          ylab = "Vertical Scaled Score")
```



#Correlation coefficient analysis for Disability Status. THIS MODEL IS ALRIGHT, BUT THERE IS A BETTER ONE IN POWERBI!!!
```{r}
ggscatter(Disability_Status,
          x = "Disability Status",
          y = "Vertical Scaled Score",
          add = "reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          cor.method = "pearson",
          xlab = "Disability Status",
          ylab = "Vertical Scaled Score")
```